Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-30 arch.tex
More edits based on Ed's comments.

Although the dataflow engine automatically propagates information about
non-null and invalid metadata, it also requires for these flows to be
initialized. That is, there needs to be some way to mark metadata as non-null
or invalid. These interfaces from the monitoring core and the dropping hardware
are marked in Figure~\ref{fig:arch.overview}.

@@

The invalid flags for array bounds check operate similarly to the null flags.
When ALU, load, or store events are dropped or filtered, the destination
register or memory location is marked as invalid. In addition, we need to
handle the possibility of dropping an array allocation. In this case, the array
pointer is marked as invalid. In terms of filtering out invalid events, for an
ALU event, if either of the source registers is invalid then the event can be
filtered. For load and store instructions, if the single source operand is
invalid, then the event can be filtered.

@@

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination.  For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a null and invalid flag with a single piece of metadata.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as null may indicate that the entire data structure's metadata
is null.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support more complex monitoring schemes.

>>>>

Although the dataflow engine handles the propagation of information about
non-null and invalid metadata, it also requires for these flows to be
initialized. That is, there needs to be some way to mark metadata as non-null
or invalid. These interfaces from the monitoring core and the dropping hardware
are marked in Figure~\ref{fig:arch.overview}.

@@

The invalid flags for array bounds check operate similarly to the null flags.
When ALU, load, or store events are dropped or filtered, the destination
register or memory location is marked as invalid. In addition, we need to
handle the possibility of dropping an array allocation. In this case, the metadata for the array
pointer is marked as invalid. In terms of filtering out invalid events, for an
ALU event, if \emph{either} of the source registers is invalid then the event can be
filtered. This differs from the policy for null flags where the event can only
filtered if \emph{both} flags are marked as null. This highlights the need to
store and track separate bits for tracking null and invalid information. For
load and store instructions, if the single source operand is invalid, then the
event can be filtered.

@@

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination. For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a null and invalid flag with a single piece of metadata.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as null may indicate that the entire data structure's metadata
is null.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support these more complex monitoring schemes.

### 2014-05-30 introduction.tex
More edits based on Ed's comments.

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
% Implementing run-time monitoring in software using binary instrumentation or
% other similar methods has been shown to introduce high overheads. For example,
% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
% slowdown \cite{lift-micro06}. 
Implementing monitors in dedicated hardware has been shown to have small
performance impacts. For example, a dedicated hardware implementation of
dynamic information flow tracking (DIFT) shows performance
overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed. It 
requires for monitors to be chosen and implemented at hardware design time.
Recent studies \cite{lba-isca08} have proposed using parallel cores in a
multi-core system to perform monitoring. This allows monitoring to be done
using parallel hardware while having the flexibility and programmability of
software. However, due to the multiple cycles it may take for software to
monitor an instruction, the overheads can easily increase execution time
several fold.
% Two optimizations
In this paper, we present a hardware architecture that uses dynamic dataflow
tracking in order to reduce the overheads of programmable parallel monitoring
by filtering out unnecessary monitoring operations and by enabling partial
monitoring.

@@

Although full monitoring coverage is always desired, trading off coverage for
reduced overheads can be very useful. For example, partial
monitoring can enable a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, partial monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

% % Adjustable overhead
% There is a limit to how effective filtering is at reducing monitoring overheads. 
% Traditionally, these overheads have been an all-or-nothing cost. If the
% designer finds that the overheads exceed their desired budget, then monitoring
% cannot be enabled.  
In this paper, we show how our dataflow-guided architecture is also able to
enable partial monitoring. This is achieved by dynamically dropping monitoring
operations when overheads exceed a specified overhead budget. We prevent false
positives from partial monitoring by using the dataflow engine to track these
dropped monitoring flows. With this architecture, we investigate different
policies for deciding which events to drop for partial monitoring. These
different policies show a trade-off between closely matching the overhead
budget and the monitoring coverage achieved.

The main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
  \item We present a hardware architecture using a dataflow tracking engine to efficiently enable filtering and partial monitoring.
  \item We investigate trade-off decisions for partial monitoring on the proposed architecture.
\end{itemize}

>>>>

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
% Implementing run-time monitoring in software using binary instrumentation or
% other similar methods has been shown to introduce high overheads. For example,
% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
% slowdown \cite{lift-micro06}. 
Implementing monitors in dedicated hardware has been shown to have small
performance impacts. For example, a dedicated hardware implementation of
dynamic information flow tracking (DIFT) shows performance
overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed. It 
requires for monitors to be chosen and implemented at hardware design time.
Recent studies \cite{lba-isca08, ruwase-spaa08, paralog-asplos10, oases-sigops09} have proposed using parallel cores in a
multi-core system to perform monitoring. This allows monitoring to be done
using parallel hardware while having the flexibility and programmability of
software. However, due to the multiple cycles it may take for software to
monitor an instruction, the overheads can easily increase execution time
several fold.
% Two optimizations
In this paper, we present a hardware architecture that uses dynamic dataflow
tracking in order to reduce the overheads of programmable parallel monitoring
by filtering out unnecessary monitoring operations and by enabling partial
monitoring.

@@

There is a limit to how effective filtering is at reducing monitoring overheads
as it only eliminates unnecessary monitoring operations. The ability
to also skip relevant monitoring operations can be used to further reduce overheads.
Although this comes at the cost of reduced monitoring coverage (i.e.,
effectiveness), this partial monitoring can still be very useful.
For example, partial
monitoring can enable a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, partial monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

% % Adjustable overhead
In this paper, we show how our dataflow-guided architecture is also able to
enable partial monitoring. This is achieved by dynamically dropping monitoring
operations when overheads exceed a specified overhead budget. We prevent false
positives from partial monitoring by using the dataflow engine to track these
dropped monitoring flows. With this architecture, we investigate different
policies for deciding which events to drop for partial monitoring. These
different policies show a trade-off between closely matching the overhead
budget and the monitoring coverage achieved.

The main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
  \item We present a hardware architecture using a dataflow tracking engine to
  efficiently enable filtering and partial monitoring. This hardware
  architecture is designed to be generally applicable to a wide-range of run-time monitoring
  techniques.
  \item We investigate interesting trade-off decisions for partial monitoring on the proposed architecture.
\end{itemize}

### 2014-05-30 related.tex
More edits based on Ed's comments.

% Sampling
Statistical sampling is a method that has been applied to various debugging
techniques in order to reduce their performance overheads. For example, Liblit
et al. sampled program runs of thousands of users in order to find bugs
\cite{liblit-pldi05}. This idea has also been applied to detecting memory
leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
other run-time monitoring techniques. The Testudo project
\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
These sampling-based techniques all limit the debugging capabilities in order
to provide lower performance overheads, similar to our architecture's goals.
However, these approaches are not able to explicitly set an overhead budget, as
our system is able to do. In addition, with the exception of Testudo, these
methods have not considered parallel monitoring. Finally, we note that the
concept of distributing the monitoring over many users and runs to increase
coverage can also be enabled using our architecture.

% Limited Monitoring
In terms of controlling overheads, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
to the one we address in this paper. However, both projects are focused on
software-based monitoring and thus their mechanisms modify the software
monitoring framework in order to enable or disable monitoring. Instead, our
architecture provides a general hardware mechanism to control overheads of
parallel monitoring architectures. In addition, these works prevented
false positives by enabling and disabling monitoring using only a source-dropping
policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring in hard real-time systems. Their architecture performs unrestricted 
dropping in order to guarantee strict deadlines. Our system allows for a
trade-off between choosing source-only dropping, unrestricted dropping, or sub-flow 
dropping depending on the needs of the designer or user which has not been
explored by these previous works.

>>>>

% Sampling
Statistical sampling is a method that has been applied to various debugging
techniques in order to reduce their performance overheads. For example, Liblit
et al. sampled program runs of thousands of users in order to find bugs
\cite{liblit-pldi05}. This idea has also been applied to detecting memory
leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
other run-time monitoring techniques. The Testudo project
\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
These sampling-based techniques all limit the debugging capabilities in order
to provide lower performance overheads, similar to our architecture's goals.
However, these approaches are not able to explicitly set an overhead budget, as
our system is able to do. In addition, with the exception of Testudo, these
methods have not considered parallel monitoring. Our techniques for generally
reducing the overheads of monitoring can also be used to enable these
low-overhead, multi-user/run debugging techniques.

% Limited Monitoring
In terms of partial monitoring, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
to the one we address in this paper. However, both projects are focused on
software-based monitoring and thus their mechanisms modify the software
monitoring framework in order to enable or disable monitoring. Instead, our
architecture provides a general hardware mechanism to control overheads of
parallel monitoring architectures. In addition, these works prevented
false positives by enabling and disabling monitoring using only a source-dropping
policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring in hard real-time systems. Their architecture performs unrestricted 
dropping in order to guarantee strict deadlines. Our system allows for a
trade-off between choosing source-only dropping, unrestricted dropping, or sub-flow 
dropping depending on the needs of the designer or user which has not been
explored by these previous works.
